---
title: "results"
execute:
  eval: false
---

Below are linear mixed model results based on our initial hypotheses around body size, thermoregulation, and daylength

Load data, bypassing the last steps of \_targets currently

```{r}
library(ggplot2)
library(inatDailyActivity)
library(lme4)
library(splines)
library(sjPlot)
library(performance)
library(dplyr)

sscs <- readRDS("_targets/objects/sscs")

daymet_data <- readRDS("_targets/objects/daymet_data")

daymet_data <- daymet_data %>% group_by(season,cell) %>% summarise(tmax = mean(tmax),daylength=mean(daylength))

sscs$cell <- as.factor(sscs$cell)
sscs$season <- as.factor(sscs$season)
daymet_data$cell <- as.factor(daymet_data$cell)
daymet_data$season <- as.factor(daymet_data$season)

sscs <- left_join(sscs, daymet_data, by = c("cell", "season"))
sscs <- mergeLeptraitsToSSCs(sscs)


sscs$midday_dist <- abs(sscs$median_value - 14)
sscs$duration <- sscs$q90_value - sscs$q10_value

sscs <- readRDS("_targets/objects/sscs_final")

sscs_scaled <- sscs %>% mutate(wingspan=scale(wingspan),tmax=scale(tmax),daylength=scale(daylength))
```

Final models and plotting

```{r}

library(car)

sscs_scaled <- sscs %>% mutate(wingspan=scale(wingspan),tmax=scale(tmax),daylength=scale(daylength))

m <- lmer(median_value ~ ns(tmax,2) + wingspan + open_closed + daylength + (1|species), data = sscs_scaled)
summary(m)
plot_model(m)
vif(m)
r2_nakagawa(m)

m <- lmer(q10_value ~ ns(tmax,2) + wingspan + open_closed + daylength + (1|species), data = sscs_scaled)
summary(m)
plot_model(m)

m <- lmer(q90_value ~ ns(tmax,2) + wingspan + open_closed + daylength + (1|species), data = sscs_scaled)
summary(m)
plot_model(m)

m <- lmer(duration ~ ns(tmax,2) + wingspan + open_closed + daylength + (1|species), data = sscs_scaled)
summary(m)
plot_model(m)
plot_model(m, type="eff", terms="daylength")

#AIC goes down with interaction effect
```

Look at just summers and lat

```{r}

#obs_data_gridded <- readRDS("_targets/objects/obs_data_gridded")
#cells <- obs_data_gridded %>% select(-geometry) %>% distinct(cell, .keep_all=TRUE) %>% select(cell_lat,cell_lon,cell) %>% mutate(cell = as.factor(cell))

summer_sscs <- filter(sscs, season == 3)
#summer_sscs <- left_join(summer_sscs, cells, by="cell")
summer_sscs_scaled <- summer_sscs %>% mutate(tmax = scale(tmax),daylength=scale(daylength))

m <- lmer(q10_value ~ ns(tmax,2) * daylength + wingspan + (1|species),data=summer_sscs_scaled)
plot_model(m)
vif(m)

m <- lmer(median_value ~ ns(tmax,2) * daylength + wingspan + (1|species),data=summer_sscs_scaled)
plot_model(m)
vif(m)

m <- lmer(q90_value ~ ns(tmax,2) * daylength + wingspan + (1|species),data=summer_sscs_scaled)
plot_model(m)
vif(m)

m <- lmer(duration ~ ns(tmax,2) * daylength + wingspan + (1|species),data=summer_sscs_scaled)
plot_model(m)
vif(m)
plot_model(m,type="eff",terms="tmax")
plot_model(m,type="eff",terms="daylength")
```

Basically, while duration increases with daylength in all seasons, in summer duration actually decreases - maybe because of the narrow range of daylength differences?

```{summary(m)}
r2_nakagawa(m)
plot_model(m, terms="tmax",type="eff")

m <- lmer(q10_value ~ tmax + (1|species),data=summer_sscs)
plot_model(m)
summary(m)
r2_nakagawa(m)
plot_model(m, terms="tmax",type="eff")

m <- lmer(q90_value ~ tmax + (1|species),data=summer_sscs)
plot_model(m)
summary(m)
r2_nakagawa(m)
plot_model(m, terms="tmax",type="eff")

ggplot(summer_sscs, aes(x=tmax)) +
  geom_density(fill = "blue", alpha = 0.5) +
  theme_minimal() +
  labs(title = "Density Plot of Continuous Variable",
       x = "Value",
       y = "Density")
```

Recode snip

```{r}

sscs <- sscs %>% mutate(season = recode(season, 
                                        '1' = 'Winter',
                                        '2' = 'Spring',
                                        '3' = 'Summer',
                                        '4' = 'Fall'))
```

Steps to arrive at these (diagnostics etc.)

```{r}

library(car)

model <- lmer(median_value ~ ns(tmax,2) * wingspan + (1|species), data = sscs)
summary(model)
plot_model(model)
plot_model(model,type="eff",terms="wingspan")
plot_model(model,type="eff",terms="tmax")
plot_model(model, type = "re")
r2_nakagawa(model)
vif(model)

model <- lmer(duration ~ tmax wingspan + (1|species), data = sscs)
summary(model)
plot_model(model)
plot_model(model,type="eff",terms="wingspan")
plot_model(model,type="eff",terms="tmax")
plot_model(model, type = "re")
r2_nakagawa(model)



model <- lmer(duration ~ tmax + wingspan + (1|species), data = sscs)
summary(model)
plot_model(model)
plot_model(model,type="eff",terms="wingspan")
plot_model(model,type="eff",terms="tmax")
plot_model(model, type = "re")
r2_nakagawa(model)

```

```{r}
library(ggplot2)

View(sscs)

# Create a data frame
data <- data.frame(hours = sscs$obs[[4]])

# Create a density plot
ggplot(data, aes(x = hours)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Butterfly Observations Across the Day", x = "Hour of the Day", y = "Density") + 
    scale_x_continuous(breaks = seq(0, 23, by = 1), labels = seq(0, 23, by = 1))

```

```{r}

# might just take more heat to warm up a large butterfly to the point they hit their peak activity 
# look into the lit about large butterfly thermregulation 

ggplot(sscs, aes(x = season, y = median_value)) +
  geom_boxplot(outlier.shape = NA) +
  labs(title = "Hour of Median Activity Across Seasons", y = "Hour of the Day", x = "Season") +
  scale_y_continuous(breaks = c(10, 12, 14, 16, 18),
                     labels = c("10 AM", "12 PM", "2 PM", "4 PM", "6 PM"),
                     limits = c(10, 18)) +
  #ylim(10,18) +
  geom_hline(yintercept = 12, color = "red", linetype = "dashed") +
  theme_minimal()

```

```{r}

plot(sscs$tmax, sscs$median_value)
```

```{r}

# Plotting the data
ggplot(sscs, aes(x=tmax, y=median_value)) +
  geom_point() +                    # scatter plot
  geom_smooth(method='loess', span = 0.75) +       # linear regression line
  labs(
    title = "Scatterplot of Value vs tmaxerature",
    x = "tmaxerature (°C)",
    y = "Value"
  ) +
  theme_minimal()

```

```{r}

# Plotting the data
ggplot(sscs, aes(x=wingspan, y=median_value)) +
  geom_point() +                    # scatter plot
  geom_smooth(method='lm') +       # linear regression line
  labs(
    title = "Scatterplot of Value vs Wingspan",
    x = "Wingspan",
    y = "Value"
  ) +
  theme_minimal()

```

```{r}

sscs_sp <- sscs %>%
  group_by(species) %>%
  tally(sort = TRUE) %>%
  top_n(20, n) %>%
  inner_join(sscs, by = "species")

ggplot(sscs_sp, aes(x = species, y = median_hr_bs)) +
  geom_boxplot(outlier.shape = NA) +
  coord_flip() + 
  labs(title = "Hour of Median Activity Across Seasons", y = "Hour of the Day", x = "Season") +
  scale_y_continuous(breaks = c(10, 12, 14, 16, 18),
                     labels = c("10 AM", "12 PM", "2 PM", "4 PM", "6 PM"),
                     limits = c(10, 18)) +
  #ylim(10,18) +
  geom_hline(yintercept = 12, color = "red", linetype = "dashed") +
  theme_minimal()
```

New interaction model

```{r}
library(lme4)
library(splines)
library(sjPlot)
library(car)
library(performance)

sscs_scaled <- sscs
sscs_scaled$tmax <- scale(sscs$tmax)
sscs_scaled$wingspan <- scale(sscs$wingspan)

#ns(tmax,2)
model <- lmer(median_value ~ ns(tmax,2) + wingspan + (1|species), data = sscs_scaled)
plot_model(model)
plot_model(model,type="eff",terms="wingspan")
plot_model(model,type="eff",terms="tmax")
plot_model(model, type = "re")
r2_nakagawa(model)
vif(model)
AIC(model)
summary(model)
library(effects)
plot(allEffects(model),rows=1)
m1 <- model


model <- lmer(q10_value ~ ns(tmax,2) + wingspan + (1|species), data = sscs_scaled)
plot_model(model)
plot_model(model,type="eff",terms="wingspan")
plot_model(model,type="eff",terms="tmax")
plot_model(model, type = "re")
r2_nakagawa(model)
vif(model)
AIC(model)
summary(model)
library(effects)
plot(allEffects(model),rows=1)
m2 <- model

model <- lmer(q90_value ~ ns(tmax,2) + wingspan + (1|species), data = sscs_scaled)
plot_model(model)
plot_model(model,type="eff",terms="wingspan")
plot_model(model,type="eff",terms="tmax")
plot_model(model, type = "re")
r2_nakagawa(model)
vif(model)
AIC(model)
summary(model)
library(effects)
plot(allEffects(model),rows=1)
m3 <- model


model <- lmer(duration ~ ns(tmax,2) + wingspan + (1|species), data = sscs_scaled)
plot_model(model)
plot_model(model,type="eff",terms="wingspan")
plot_model(model,type="eff",terms="tmax")
plot_model(model, type = "re")
r2_nakagawa(model)
vif(model)
AIC(model)
summary(model)
library(effects)
plot(allEffects(model),rows=1)
m4 <- model
```

Checking diff combos

```{r}

print("Spline, Interaction")
model <- lmer(median_value ~ ns(tmax,2) * wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("No Spline, Interaction")
model <- lmer(median_value ~ tmax * wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("Spline, No Interaction")
model <- lmer(median_value ~ ns(tmax,2) + wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("No Spline, No Interaction")
model <- lmer(median_value ~ tmax + wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)
```

```{r}

print("Spline, Interaction")
model <- lmer(q10_value ~ ns(tmax,2) * wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("No Spline, Interaction")
model <- lmer(q10_value  ~ tmax * wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("Spline, No Interaction")
model <- lmer(q10_value ~ ns(tmax,2) + wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("No Spline, No Interaction")
model <- lmer(q10_value ~ tmax + wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)
```

```{r}

print("Spline, Interaction")
model <- lmer(q90_value ~ ns(tmax,2) * wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("No Spline, Interaction")
model <- lmer(q90_value ~ tmax * wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("Spline, No Interaction")
model <- lmer(q90_value ~ ns(tmax,2) + wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("No Spline, No Interaction")
model <- lmer(q90_value ~ tmax + wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)
```

```{r}

print("Spline, Interaction")
model <- lmer(midday_dist ~ ns(tmax,2) * wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("No Spline, Interaction")
model <- lmer(midday_dist ~ tmax * wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("Spline, No Interaction")
model <- lmer(midday_dist ~ ns(tmax,2) + wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)

print("No Spline, No Interaction")
model <- lmer(midday_dist ~ tmax + wingspan + (1|species), data = sscs_scaled)
r2_nakagawa(model)
AIC(model)
```

```{r}


```

```{r}

model <- lmer(median_value ~ ns(tmax,2) * season + wingspan + (1|species), data = sscs_scaled)
plot_model(model)
plot_model(model,type="eff",terms=c("tmax","season"))
vif(model)
```
